---
title: "NZFT_SQ_pop-proj_v4"
author: "T. McMurdo Hamilton"
date: "25/11/2019"
output:
  word_document: default
  pdf_document: default
  html_document: default
---
  
```{r setup, include=FALSE}

library(popbio)
library(mc2d)
library(sjmisc)
library(tidyr)
library(plyr)
library(reshape2) 
library(dplyr)
library(ggplot2)
library(lme4)
#library(Hmisc)
library(DHARMa)
library(MuMIn)
library(lattice)
library(sjPlot)
library(plotrix)
library(truncnorm)
library(gridExtra)


```

# Population projection model with parametric uncertainty, environmental and demographic stochasticity

Transition rates in this matrix have been calculated from RMark CJS models (juvenile, immature and adult survival rates) and breeding productivity GLMMs (probability of an egg hatching (unmanaged clutch and managed clutch), fertile eggs only).

Other variables used have been calculated from the NZFT databases. Proportion of age 2+ females breeding and mean proportion of fertile nests managed are calculated from data spanning 2007-2017, mean clutch size is taken from all eggs and clutches, excluding 4th clutches, 1997-2017. Proportion of infertile eggs has been calculated from data spanning 1997-2017.

Mean clutch size, all clutches (excluding 4th) (1997-2017):
  n     mean        sd         se lower.ci upper.ci
289 1.730104 0.4446757 0.02615739  1.67862 1.781588

Mean fertility rate of all eggs laid per season (1997-2017):
  0.679 (sd = 0.156)

Mean proportion of nests managed per season (calculated from 2007-2017 to represent more recent practices):
  0.542 (sd = 0.21, se = 0.063)

Mean proportion of females (aged 2+) attempting to breed (1997-2017):
  0.665 (sd = 0.18)
This statistic used since if we do not know underlying causes of females breeding and it seems quite random, then using the biggest dataset or equivalent date span to the ones above makes more sense and is the most conservative stance. This also limits uncertainty. 

These statistics are also available:
  Mean proportion of females (aged 2+) attempting to breed (1993-2017):
  0.606 (sd = 0.21)

Mean proportion of females (aged 2+) attempting to breed (2007-2017):
  0.72 (sd = 0.12)
  
# Changes to previous code #

The following changes have been made compared to previous versions. There will ultimately be a fourth change, when I allow impact of management (X) to vary also.

1. *Carrying capacity* has been altered to affect the number of females that are able to breed, so once number of adult females reaches Car, no more can breed - this is the strucutre for all subsequent models too:
N[1,t+1,s] <- rpois(1, (min(Car, N[3,t,s]) * sa[t] * Fa))

2. *Clutch size* now varies with a distribution around the mean, standardising this across all projections except for ones with clutch size elicited

3. *Proportion managed* now varies with a distribution around the mean, standardising this across all projections except for ones with management effort elicited

# Uncertainty #

1. *Parametric uncertainty* was taken as: standard error from RMark output (survival), standard error from GLMMs (hatch and fledge probability), standard deviation of the mean (proportion of females attempting to breed, clutch size). Parameters are drawn from a probability distribution.

2. *Environmental or temporal variation* for survival is unknown/not calculated, so a low, generic value is given (0.01). For breeding parameters, the standard deviation of the random effects in each of the hatch and fledge GLMMs output is used. Parameters are drawn from a probability distribution, with the mean from step one. 

3. *Demographic stochasticity* is introduced by drawing values from Poisson distribution (fecundity) or Binomial (survival), using the simulated mean from step two, at each time step. 

In 2017 (latest data I have), there were 14 females and 26 males aged 2+, 3 female and 2 male aged 1 year. This will be used as the starting vector. 

## TARA-ITI: three age classes 

code derived from Kery & Schaub IPM book. 


```{r}


# Function to translate mean and sd of survival into parameters of a beta distribution

beta.params <- function(x_bar, sd.x){
  u <- x_bar*(1-x_bar)/sd.x^2-1
  alpha <- x_bar*u
  beta <- (1-x_bar)*u
  return(list(alpha = alpha, beta = beta))
}

# Define the number of years with predictions and the Monte Carlo setting

nyears <- 50 # Number of years (projection time frame)
nsim <- 10000 # Number of replicate populations simulated


# Define population matrix and initial stage-specific population sizes

N <- array(0, dim = c(3, nyears + 1, nsim)) # edited for three life stages
Ni <- c(0, 3, 14) # starting population vector

N[,1,] <- Ni 

alive <- matrix(0, nrow = nyears, ncol = nsim)
r.sq <- matrix(0, nrow = nyears, ncol = nsim)
mean.r <- lambda.sq <- numeric()
sj.val <- si.val <- sa.val <- Fa.val <- persist<- numeric()



# Define mean, measurement error and temporal variability of the demographic parameters
# NB temporal variability for survival has been generically assigned as 0.01 - this is *not* based on real data


Sjuv <- 0.8053
Sjuv.e <- 0.097                # Uncertainty of juv. survival expressed as std error from RMark output
Sjuv.t <- 0.00                 # temporal variability of juvenile survival (would be expressed as SD on logit scale)
Simm <- 0.927
Simm.e <- 0.061                # Uncertainty of imm survival expressed as std error from RMark output
Simm.t <- 0.00                 # temporal variability of immature survival (would be expressed as SD on logit scale)
Sad <- 0.919
Sad.e <- 0.023                 # Uncertainty of imm survival expressed as std error from RMark output
Sad.t <- 0.00                  # temporal variability of immature survival (would be expressed as SD on logit scale)

h <- qlogis(0.809)              # untransformed prob of eggs hatching (backtransformed from ggpredict output)
h.e <- 0.346                    # uncertainty expressed as untransformed std error from ggpredict 
h.t <- 0.2046                   # temporal variability of hatch prob (from model m2) expressed as untransformed SD in the random effects for season start var
fl <-qlogis(0.704)              # untransformed prob of hatchlings fledging (backtransformed from ggpredict output)
fl.e <-0.410                    # uncertainty expressed as untransformed std error from ggpredict 
fl.t <- 0.2389                  # temporal variability of fledge prob (from model f3) expressed as untransformed SD
K<-0.665                         # mean prop of females breeding
K.e <-0.18                      # uncertainty expressed as standard deviation of the mean
# K<-0.45                       # Ferreira et al, 2005
c <- 1.73                      # mean clutch size 
c.e <- 0.45                    # uncertainty expressed as standard deviation of the mean
M <-0.542                        # mean proportion of eggs managed (2007-2017)
M.e <- 0.21                      # uncertainty expressed as SD of the mean

## constants 

X <- 0.47                       #'cost' of management-reduction in prob of fledging from egg
# X.e <-                        # error from cost of management still to be calculated
infer <-0.679                   # mean proportion of fertile eggs 
infer.e <- 0.156                # uncertainty expressed as SD of the mean

## set a carrying capacity
 
#Car <- 100
#Car <- 22                       # highest mean value from group for field 1
Car <- 17                       # mean 'most likely' mode estimate for field 1
#Car <-37

```

Project the population:
  
```{r}


# Project population

for (s in 1:nsim){                                          # Loop over replicate populations
  if(s %% 100 == 0) {cat(paste("*** Simrep", s, "***\n")) } # Counter
  
  # Generate mean demographic values from beta and normal distributions (subject to measurement error)
  
  
  sj.sim <- rbeta(1, beta.params(Sjuv, Sjuv.e)$alpha, beta.params(Sjuv, Sjuv.e)$beta)
  si.sim <- rbeta(1, beta.params(Simm, Simm.e)$alpha, beta.params(Simm, Simm.e)$beta)
  sa.sim <- rbeta(1, beta.params(Sad, Sad.e)$alpha, beta.params(Sad, Sad.e)$beta)
  h.sim <-  rnorm(1, h, h.e)             # transformed in the next param simulation 
  fl.sim <- rnorm(1, fl, fl.e)           # transformed in the next param simulation 
  K.sim <- rbeta(1, beta.params(K, K.e)$alpha, beta.params(K, K.e)$beta)
  c.sim <- rtruncnorm(1, a=0, b=2, mean=c, sd=c.e) 
  M.sim <-  rbeta(1, beta.params(M, M.e)$alpha, beta.params(M, M.e)$beta) #remove this if removing management
  
  # Generate annual demographic rates (subject to temporal variability)
  sj <- plogis(rnorm(nyears, qlogis(sj.sim), Sjuv.t))
  si <- plogis(rnorm(nyears, qlogis(si.sim), Simm.t))
  sa <- plogis(rnorm(nyears, qlogis(sa.sim), Sad.t))
  ha <- plogis(rnorm(nyears, h.sim, h.t))
  fla <- plogis(rnorm(nyears, fl.sim, fl.t))
  Ka <- rnorm(nyears, K.sim, 0) # no temporal var in K
  ca <- rnorm(nyears, c.sim, 0)
  Ma <- rnorm(nyears, M.sim, 0) # no temporal var in M, remove this if removing management (to test effect)
  
  # Project population (include demographic stochasticity)
  
  for (t in 1:nyears){ # Loop over years
    
    Fa <- (Ka[t]*ca[t]*infer*((Ma[t]*X)+(1-Ma[t]))*(ha[t]*fla[t]))/2  # fecundity equation 
    #Fa <- (Ka[t]*ca[t]*infer*(ha[t]*fla[t]))/2  # fecundity equation without management
    N[1,t+1,s] <- rpois(1, min(Car, N[3,t,s]) * sa[t] * Fa) # carrying capacity
    N[2,t+1,s] <- rbinom(1, (N[1,t,s]), sj[t])
    N[3,t+1,s] <- rbinom(1, (N[2,t,s]), si[t]) + rbinom(1, (N[3,t,s]), sa[t])
    if (sum(N[,t+1,s]) == 0) break
    else {
      r.sq[t,s] <- log(sum(N[,t+1,s])) - log(sum(N[,t,s]))
      alive[t,s] <- t
    } # else
    
    persist[t]<- sum(N[3,t,]>2)  / s # 
    
  } # t
  
  sj.val[s] = sj.sim
  si.val[s] = si.sim
  sa.val[s] = sa.sim
  Fa.val[s] = Fa
  
  
  mean.r[s] <- mean(r.sq[min(alive[,s], na.rm = TRUE):max(alive[,s], na.rm = TRUE),s])
  lambda.sq[s] <- exp(mean.r[s]) 
  
} # s

```

# Intrinsic growth rates and population extinction probability

Calculate the mean intrinsic growth rate (plus standard deviation) of all populations, and only populations that don't go extinct. 

Calculate the probability of extinction after 50 years. Important to define what we class as 'extinction' here. The population has historically gone down to 3 known pairs. Makes sense to look at the rate of populations going to zero adults, and the rate of populations going down to 2 adult females.

Mean intrinsic growth rate (and standard deviation)

```{r}

mean(mean.r) 
sd(mean.r)

mean.l.noman<-mean(lambda.sq)
mean(lambda.sq)
sd(lambda.sq)

# store lambda

# write.csv(lambda.sq, "Lambda_SQ.08_12.csv")
```

Mean intrinsic growth rate of surviving populations
```{r}
not.extinct <- which((alive[nyears,])!=0)
mean(mean.r[not.extinct]) 
sd(mean.r[not.extinct])
```
   
Persistance probability (after nyears)

```{r}

sum(alive[nyears,]!=0) / nsim 

# #store persistence 
# 
# write.csv(persist, "persist_SQ_08_12.csv")
# 

```

Extinction probability (population goes to zero, probability goes to 2 adult females or less)

```{r}

# calculation of quasi-extinction (Stefanos code)

sum(apply(N[3,,1:nsim],2,min)<3)/nsim # adults only

sum(apply(colSums(N[-1,,1:nsim]),2,min)<3)/nsim # immatures and adults only


```

```{r}
# Quasi-extinction

sum(N[3,nyears,]<=2) / nsim
sum(N[3,nyears,]==0) / nsim
min(colSums(N[,,1:10]))<20 
```

Probability that population is equal to or smaller than the initial adult population size in year 1 (including extinct pops), and probability that population is surviving, but equal to or less than the 2017 size.

```{r}

sum(N[3,nyears,]<=N[3,1,1]) / nsim
sum((N[3,nyears,]<=14) & ( N[3,nyears,]>2)) / nsim

```

Probability that the population is persisting and is larger than the initial adult population in year 1
```{r}

sum((N[3,nyears,] > N[3,1,1])) / nsim
summary(N[3,51,])

```



## Plots of simulated populations

Persistence is identified as two adult females or fewer. This is because the lowest recorded population size was three known pairs. 

```{R}

x<-1:nyears
plot(x, persist, type = "l", ylim = c(0,1))


```


```{r}

# prep data for plots

N.mean <- apply(N,c(1,2),mean)
N.sd <- apply(N, c(1,2), sd)
N.lci <- apply(N,c(1,2),quantile, probs=c(.025))
N.uci <- apply(N,c(1,2),quantile, probs=c(.975))
N.upr <- N.mean + 1.96*N.sd                       # upper 95% confidence level (1.96 * SD)
N.lwr <- N.mean - 1.96*N.sd                       # lower 95% confidence level
N.25 <- apply(N,c(1,2),quantile, probs=c(.25))    # lower quartile
N.50 <- apply(N,c(1,2),quantile, probs=c(.5))     # median quartile
N.75 <- apply(N,c(1,2),quantile, probs=c(.75))    # upper quartile


plot_results = data.frame(alt = "SQ", years = 1:51, N.mean=N.mean[3,], N.uci = N.uci[3,], N.lci=N.lci[3,], N.upr = N.upr[3,], N.lwr = N.lwr[3,],
                          N.25 = N.25[3,],N.50 = N.50[3,], N.75 = N.75[3,])

# # store projections
# 
# write.csv(plot_results, "plot_results_SQ_08_12.csv")

gg2<-ggplot(plot_results, aes(x=years, y=N.mean)) + 
 geom_ribbon(aes(x = years, ymax=N.upr, ymin=N.lwr),fill = "lightgrey", alpha=0.35) +   geom_ribbon(aes(x = years, ymax=N.75, ymin=N.25), fill = "dimgrey", alpha=0.35) + geom_line(aes(x = years, y=N.50), color="white", size = 2, alpha=0.6) + geom_line(color="black", size=1, alpha=0.6) + theme_classic(base_size = 16)+  coord_cartesian(ylim=c(0,max(N.upr+10)))  + labs(title = "Predicted mean tara iti population size", x = "Year", y = "Population size (adult females)")

gg2


## store the population sizes at the end of the simulation

N_50_sq<-N[3,50,1:nsim]
N_25_sq<-N[3,25,1:nsim]
# 
# write.csv(N_50_sq, "Finalpopsize_SQ_08_12.csv")

# Plot graph with population growth rates and population size

layout(matrix(c(1,1,2,3), 2, 2, byrow = TRUE))
par(mfrow=c(1,1))
par(las = 1, cex = 1.1)
plot(r.sq[,1], type = "l", lwd = 0.5, ylab = "Annual population growth rate", xlab = "Time", ylim = range(r.sq[which(!is.na(alive))]), col = "lightgrey")
for (s in 2:nsim){
  lines(r.sq[!is.na(alive[,s]),s], lwd = 0.5, col = "lightgrey")
}

lines(apply(r.sq, 1, mean, na.rm = TRUE), lwd = 1.5)
legend("bottomright", lwd = c(1,1), col = c("lightgrey", "black"), legend = c("Individual", "Mean"), bty = "n")
text(x = 0, y = 1.3, "a")

a <- hist(mean.r, nclass = 25, col = "dodgerblue", main = "", xlab = "Population growth rate")
text(x = a$mids[1], y = max(a$counts)*0.95, "b")
box()

a <- hist(mean.r[not.extinct], nclass = 25, col = "dodgerblue", main = "", xlab = "Population growth rate")
text(x = a$mids[1], y = max(a$counts)*0.95, "c")
box()

par(mfrow=c(1,1))

plot(N[1,,s], type = "l", lwd = 0.5, ylab = "Population size", xlab = "Time", ylim = c(0,200), col = "white")
for (s in 2:nsim){
  lines(N[3,,s], lwd = 0.5, col = "dodgerblue") # plot just the adults
  #lines(colSums(N[,,s]), lwd = 0.5, col = "dodgerblue") # if plotting whole pop
}


hist(N[3,50,1:nsim], col="dodgerblue", xlim=c(0,100),
     breaks=seq(0,1000, 5), xlab="Final population size at t=50", main='')


gg.sq<-ggplot(plot_results, aes(x=years, y=N.mean)) + 
geom_ribbon(aes(x = years, ymax=N.upr, ymin=N.lwr),fill = "dimgrey", alpha=0.35) + geom_line(color="mediumblue", size=1.5, alpha=0.6) + theme_classic(base_size = 16)+  coord_cartesian(ylim=c(0,max(N.upr+10)))  + labs(title = "Predicted mean tara iti population size", x = "Year", y = "Population size (adult females)")

gg.sq

#plot(density(SQ_N_50), xlab="Final population size at t=50", main='', lwd = 2) #+ lines(density(SQ_N_25))

lambda.density<-plot(density(lambda.sq), lwd = 3, xlim=c(0.8, 1.2), main = "",  xlab = expression(lambda)) + abline(v=1, col = "grey")


```


# Correlation of vital rates with mean intrinsic growth rate

```{R}

par(mfrow=c(2,2))

corr_df<-data.frame(sj<-sj.val, si<-si.val, sa<- sa.val, Fa<-Fa.val, r<-mean.r)

cor2<-ggplot(corr_df, aes(sj, r)) + geom_point(shape = 21, size = 1.5) + labs(title = "Correlation of juvenile survival + mean population growth rate") +   theme_classic(base_size = 12) + labs(title = "Correlation of juvenile survival + mean population growth rate") +
  theme_classic(base_size = 12)

cor3<-ggplot(corr_df, aes(si, r)) + geom_point(shape = 21, size = 1.5) + labs(title = "Correlation of juvenile survival + mean population growth rate") +   theme_classic(base_size = 12) + labs(title = "Correlation of immature survival + mean population growth rate") +
  theme_classic(base_size = 12)

cor4<-ggplot(corr_df, aes(sa, r)) + geom_point(shape = 21, size = 1.5) + labs(title = "Correlation of juvenile survival + mean population growth rate") +   theme_classic(base_size = 12) + labs(title = "Correlation of adult survival + mean population growth rate") +
  theme_classic(base_size = 12)

cor5<-ggplot(corr_df, aes(Fa, r)) + geom_point(shape = 21, size = 1.5) + labs(title = "Correlation of juvenile survival + mean population growth rate") +   theme_classic(base_size = 12) + labs(title = "Correlation of fecundity + mean population growth rate") +
  theme_classic(base_size = 12)

grid.arrange(cor2, cor3, cor4, cor5, ncol=2)
```
